\bf Approximating Likelihood Ratios with Calibrated Discriminative Classifiers

### 2016-01-21 jasa/paper-jasa.tex

Let denote the maximum likelihood estimator
\begin{equation}\label{eq:mle}
\hat{\theta} = \argmax_\theta  p({\cal D} | \theta) \; .
\end{equation}
% and the conditional maximum likelihood estimator
% \begin{equation}\label{eq:cmle}
% \hat{\hat{\theta}} = \argmax_{\theta \in \Theta_0}  p({\cal D} | \theta) \; .
% \end{equation}
It is not obvious that if we are working instead with distributions $p(s(\mathbf{x})|\theta)$
that we can find the
same estimators. Fortunately, there is a construction based on $s$
that works. The maximum likelihood estimate of Eqn.~\ref{eq:mle} is the same as
the value that maximizes the likelihood ratio with respect to $p({\cal
D}|\theta_1)$ for some fixed value of $\theta_1$. This allows us to use
Theorem~\ref{thm:ratio-equivalence} to reformulate the maximum likelihood estimate:
\begin{align}\label{eq:mle_withs}
\hat{\theta} &= \argmax_\theta  p({\cal D} | \theta) \nonumber \\
&= \argmax_\theta  \prod \frac{p(\mathbf{x}| \theta)}{p(\mathbf{x}|\theta_1)} \nonumber \\
&= \argmax_\theta  \prod \frac{p(s(\mathbf{x}; \theta, \theta_1) | \theta)}{p(s(\mathbf{x}; \theta, \theta_1) |\theta_1)} \;
\end{align}
where $s(\mathbf{x};\theta,\theta_1)$ denotes a \textit{parametrized} transformation $s$
of $X$ in terms of $(\theta,\theta_1)$. Note that it is important that we include the denominator $p(s(\mathbf{x}; \theta, \theta_1)
|\theta_1)$ because this cancels Jacobian factors that vary with $\theta$.

>>>>

Evaluating Eqn.~\ref{eqn:generalized-lr} requires finding the maximum likelihood
estimator
\begin{equation}\label{eq:mle}
    \hat{\theta} = \argmax_\theta p({\cal D} | \theta),
\end{equation}
which is again
made difficult in the likelihood-free setting.  Also, it is not obvious that if
we are working instead with distributions $p(s(\mathbf{x})|\theta)$ that we can
find the same estimators. Fortunately, there is a construction based on $s$ that
works: the maximum likelihood estimate of Eqn.~\ref{eq:mle} is the same as the
value that maximizes the likelihood ratio with respect to $p({\cal D}|\theta_1)$,
for some fixed value of $\theta_1$. This allows us to use
Theorem~\ref{thm:ratio-equivalence} to reformulate the maximum likelihood
estimate as
\begin{align}\label{eq:mle_withs}
\hat{\theta} &= \argmax_\theta  p({\cal D} | \theta) \nonumber \\
&= \argmax_\theta  \prod \frac{p(\mathbf{x}| \theta)}{p(\mathbf{x}|\theta_1)} \nonumber \\
&= \argmax_\theta  \prod \frac{p(s(\mathbf{x}; \theta, \theta_1) | \theta)}{p(s(\mathbf{x}; \theta, \theta_1) |\theta_1)} \;,
\end{align}
where $s(\mathbf{x};\theta,\theta_1)$ denotes a \textit{parametrized}
transformation $s$ of $X$ in terms of $(\theta,\theta_1)$ that is monotonic
with $r(\mathbf{x};\theta,\theta_1)$. Note that it is important that we include
the denominator $p(s(\mathbf{x}; \theta, \theta_1) |\theta_1)$ because this
cancels Jacobian factors that vary with $\theta$.
